---
title: "Brightness Analysis: Edge Detection and Surface Contamination"
author: "Brandon Titensor"
date: "2025-06-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Introduction

This analysis examines brightness measurements from star shade contamination trials, including:
- **Edge measurements**: Star shade and calibration edge detection (upper/lower regions)
- **Surface measurements**: Contamination detection across multiple surface samples
- **Sample types**: Star shade surfaces, calibration wafer surfaces, and witness samples

# Data Loading and Setup

```{r Data}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(knitr)
library(corrplot)
library(stringr)

# Load all brightness datasets
star_edge <- read_csv("/Volumes/BRANDONMEGA/Research/Dust_Contamination/Trials/Test_Files/Brightness/Edge/Median_Intensity_Results.csv")
cal_edge <- read_csv("/Volumes/BRANDONMEGA/Research/Dust_Contamination/Trials/Test_Files/Brightness/Calibration_Edge/Median_Intensity_Results.csv")
surfaces <- read_csv("/Volumes/BRANDONMEGA/Research/Dust_Contamination/Trials/Test_Files/Brightness/Surfaces/Surface_Median_Intensity_Results.csv")

# Display structure of each dataset
cat("Star Edge Data:\n")
head(star_edge, 3)
cat("\nCalibration Edge Data:\n")
head(cal_edge, 3)
cat("\nSurface Data:\n")
head(surfaces, 3)
```

# Data Cleaning and Preparation

```{r data_cleaning}
# Clean edge datasets (star_edge and cal_edge)
colnames(star_edge) <- c("Folder_Name", "File_Name", "Upper_40_Median", 
                        "Upper_30_Median", "Lower_30_Median", "Entire_Image_Median")
colnames(cal_edge) <- c("Folder_Name", "File_Name", "Upper_40_Median", 
                       "Upper_30_Median", "Lower_30_Median", "Entire_Image_Median")

# Clean surface dataset
colnames(surfaces) <- c("Folder_Name", "File_Name", "Entire_Image_Median",
                       "Sample1_Median", "Sample2_Median", "Sample3_Median",
                       "Sample4_Median", "Sample5_Median", "Sample6_Median", "Sample7_Median")

# Function to extract trial and sample information from folder names
extract_trial_info <- function(folder_name) {
  folder_clean <- gsub("/", "", folder_name)
  
  # Extract trial number (looking for patterns like "Tr10", "Tr12", etc.)
  trial_match <- str_extract(folder_clean, "Tr\\d+")
  trial <- ifelse(is.na(trial_match), "Unknown", trial_match)
  
  # Extract sample number (looking for patterns like "Sa1", "Sa2", etc.)
  sample_match <- str_extract(folder_clean, "Sa\\d+")
  sample <- ifelse(is.na(sample_match), "Unknown", sample_match)
  
  # Determine timing (before/after)
  timing <- case_when(
    str_detect(folder_clean, "^Bef") ~ "Before",
    str_detect(folder_clean, "^Aft") ~ "After", 
    str_detect(folder_clean, "Witness_bef") ~ "Before",
    str_detect(folder_clean, "Witness_aft") ~ "After",
    TRUE ~ "Unknown"
  )
  
  # Determine surface type
  surface_type <- case_when(
    str_detect(folder_clean, "Surf") ~ "Surface",
    str_detect(folder_clean, "CwSurf") ~ "Calibration_Wafer_Surface",
    str_detect(folder_clean, "Witness") ~ "Witness",
    TRUE ~ "Edge"
  )
  
  return(list(trial = trial, sample = sample, timing = timing, surface_type = surface_type))
}

# Add dataset identifiers and process edge data
star_edge <- star_edge %>%
  mutate(
    Dataset = "Star_Edge",
    Image_Number = as.numeric(gsub(".*?(\\d+)\\.jpg$", "\\1", File_Name)),
    Folder_Clean = gsub("/", "", Folder_Name)
  ) %>%
  rowwise() %>%
  mutate(
    trial_info = list(extract_trial_info(Folder_Name)),
    Trial = trial_info$trial,
    Sample = trial_info$sample, 
    Timing = trial_info$timing,
    Surface_Type = trial_info$surface_type
  ) %>%
  select(-trial_info) %>%
  ungroup() %>%
  arrange(Trial, Sample, Image_Number)

cal_edge <- cal_edge %>%
  mutate(
    Dataset = "Calibration_Edge", 
    Image_Number = as.numeric(gsub(".*?(\\d+)\\.jpg$", "\\1", File_Name)),
    Folder_Clean = gsub("/", "", Folder_Name)
  ) %>%
  rowwise() %>%
  mutate(
    trial_info = list(extract_trial_info(Folder_Name)),
    Trial = trial_info$trial,
    Sample = trial_info$sample,
    Timing = trial_info$timing,
    Surface_Type = trial_info$surface_type
  ) %>%
  select(-trial_info) %>%
  ungroup() %>%
  arrange(Trial, Sample, Image_Number)

# Process surface data and categorize sample types
surfaces <- surfaces %>%
  mutate(
    Dataset = "Surface",
    Image_Number = as.numeric(gsub(".*?(\\d+)\\.jpg$", "\\1", File_Name)),
    Folder_Clean = gsub("/", "", Folder_Name),
    Sample_Type = case_when(
      str_detect(File_Name, "starshade") ~ "Star_Shade_Surface",
      str_detect(File_Name, "cali") ~ "Calibration_Surface", 
      str_detect(File_Name, "witness") ~ "Witness_Sample",
      TRUE ~ "Other"
    )
  ) %>%
  rowwise() %>%
  mutate(
    trial_info = list(extract_trial_info(Folder_Name)),
    Trial = trial_info$trial,
    Sample = trial_info$sample,
    Timing = trial_info$timing,
    Surface_Type = trial_info$surface_type
  ) %>%
  select(-trial_info) %>%
  ungroup() %>%
  arrange(Trial, Sample, Image_Number)

# Display cleaned data summaries
cat("Dataset sizes:\n")
cat("Star Edge:", nrow(star_edge), "images\n")
cat("Calibration Edge:", nrow(cal_edge), "images\n") 
cat("Surfaces:", nrow(surfaces), "images\n")

cat("\nTrial and Sample breakdown:\n")
cat("Star Edge Trials:", paste(unique(star_edge$Trial), collapse = ", "), "\n")
cat("Star Edge Samples:", paste(unique(star_edge$Sample), collapse = ", "), "\n")
cat("Cal Edge Trials:", paste(unique(cal_edge$Trial), collapse = ", "), "\n")
cat("Cal Edge Samples:", paste(unique(cal_edge$Sample), collapse = ", "), "\n")
cat("Surface Trials:", paste(unique(surfaces$Trial), collapse = ", "), "\n")
cat("Surface Samples:", paste(unique(surfaces$Sample), collapse = ", "), "\n")
```

# Descriptive Statistics by Trial and Sample

## Edge Measurements by Trial and Sample

```{r edge_descriptive_stats}
# Star Edge descriptive statistics
star_edge_stats <- star_edge %>%
  group_by(Trial, Sample, Timing, Surface_Type) %>%
  summarise(
    N_Images = n(),
    Upper_40_Mean = round(mean(Upper_40_Median, na.rm = TRUE), 2),
    Upper_40_SD = round(sd(Upper_40_Median, na.rm = TRUE), 2),
    Upper_30_Mean = round(mean(Upper_30_Median, na.rm = TRUE), 2), 
    Upper_30_SD = round(sd(Upper_30_Median, na.rm = TRUE), 2),
    Lower_30_Mean = round(mean(Lower_30_Median, na.rm = TRUE), 2),
    Lower_30_SD = round(sd(Lower_30_Median, na.rm = TRUE), 2),
    Entire_Mean = round(mean(Entire_Image_Median, na.rm = TRUE), 2),
    Entire_SD = round(sd(Entire_Image_Median, na.rm = TRUE), 2),
    Entire_Min = min(Entire_Image_Median, na.rm = TRUE),
    Entire_Max = max(Entire_Image_Median, na.rm = TRUE),
    .groups = 'drop'
  )

kable(star_edge_stats, 
      caption = "Star Edge Descriptive Statistics by Trial and Sample")

# Calibration Edge descriptive statistics
cal_edge_stats <- cal_edge %>%
  group_by(Trial, Sample, Timing, Surface_Type) %>%
  summarise(
    N_Images = n(),
    Upper_40_Mean = round(mean(Upper_40_Median, na.rm = TRUE), 2),
    Upper_40_SD = round(sd(Upper_40_Median, na.rm = TRUE), 2),
    Upper_30_Mean = round(mean(Upper_30_Median, na.rm = TRUE), 2),
    Upper_30_SD = round(sd(Upper_30_Median, na.rm = TRUE), 2),
    Lower_30_Mean = round(mean(Lower_30_Median, na.rm = TRUE), 2),
    Lower_30_SD = round(sd(Lower_30_Median, na.rm = TRUE), 2),
    Entire_Mean = round(mean(Entire_Image_Median, na.rm = TRUE), 2),
    Entire_SD = round(sd(Entire_Image_Median, na.rm = TRUE), 2),
    Entire_Min = min(Entire_Image_Median, na.rm = TRUE),
    Entire_Max = max(Entire_Image_Median, na.rm = TRUE),
    .groups = 'drop'
  )

kable(cal_edge_stats,
      caption = "Calibration Edge Descriptive Statistics by Trial and Sample")
```

## Surface Measurements by Trial and Sample

```{r surface_descriptive_stats}
# Calculate total contamination and descriptive stats for surfaces
surfaces_with_totals <- surfaces %>%
  mutate(
    Total_Contamination = Sample1_Median + Sample2_Median + Sample3_Median + 
                         Sample4_Median + Sample5_Median + Sample6_Median + Sample7_Median,
    Max_Sample_Contamination = pmax(Sample1_Median, Sample2_Median, Sample3_Median,
                                   Sample4_Median, Sample5_Median, Sample6_Median, Sample7_Median),
    Contaminated_Samples = (Sample1_Median > 0) + (Sample2_Median > 0) + (Sample3_Median > 0) +
                          (Sample4_Median > 0) + (Sample5_Median > 0) + (Sample6_Median > 0) + (Sample7_Median > 0),
    Clean_Images = Total_Contamination == 0
  )

# Surface descriptive statistics
surface_stats <- surfaces_with_totals %>%
  group_by(Trial, Sample, Timing, Surface_Type, Sample_Type) %>%
  summarise(
    N_Images = n(),
    Entire_Image_Mean = round(mean(Entire_Image_Median, na.rm = TRUE), 2),
    Entire_Image_SD = round(sd(Entire_Image_Median, na.rm = TRUE), 2),
    Total_Contamination_Mean = round(mean(Total_Contamination, na.rm = TRUE), 2),
    Total_Contamination_SD = round(sd(Total_Contamination, na.rm = TRUE), 2),
    Max_Contamination_Mean = round(mean(Max_Sample_Contamination, na.rm = TRUE), 2),
    Max_Contamination_SD = round(sd(Max_Sample_Contamination, na.rm = TRUE), 2),
    Avg_Contaminated_Samples = round(mean(Contaminated_Samples, na.rm = TRUE), 2),
    Clean_Rate_Percent = round(sum(Clean_Images) / n() * 100, 1),
    .groups = 'drop'
  )

kable(surface_stats,
      caption = "Surface Contamination Descriptive Statistics by Trial and Sample")

# Individual sample position statistics by trial and sample
sample_position_stats <- surfaces %>%
  select(Trial, Sample, Timing, Surface_Type, Sample_Type, Sample1_Median:Sample7_Median) %>%
  pivot_longer(cols = Sample1_Median:Sample7_Median, 
               names_to = "Sample_Position", values_to = "Contamination_Level") %>%
  mutate(Sample_Position = gsub("_Median", "", Sample_Position)) %>%
  group_by(Trial, Sample, Timing, Surface_Type, Sample_Type, Sample_Position) %>%
  summarise(
    N_Measurements = n(),
    Mean_Contamination = round(mean(Contamination_Level, na.rm = TRUE), 2),
    SD_Contamination = round(sd(Contamination_Level, na.rm = TRUE), 2),
    Max_Contamination = max(Contamination_Level, na.rm = TRUE),
    Contamination_Events = sum(Contamination_Level > 0, na.rm = TRUE),
    Contamination_Rate = round(sum(Contamination_Level > 0, na.rm = TRUE) / n() * 100, 1),
    .groups = 'drop'
  )

# Show sample position stats in a more readable format
sample_pos_summary <- sample_position_stats %>%
  select(Trial, Sample, Sample_Type, Sample_Position, Mean_Contamination, Contamination_Rate) %>%
  pivot_wider(names_from = Sample_Position, 
              values_from = c(Mean_Contamination, Contamination_Rate),
              names_sep = "_")

kable(sample_pos_summary,
      caption = "Sample Position Statistics by Trial and Sample (Mean Contamination and Rate %)")
```

# Edge Analysis: Star Shade vs Calibration

## Trial and Sample Comparison Visualization

```{r trial_sample_visualization}
# Edge measurements comparison by trial and sample
edge_combined <- bind_rows(star_edge, cal_edge)

# Box plot of entire image median by trial and sample
ggplot(edge_combined, aes(x = paste(Trial, Sample, sep = "-"), y = Entire_Image_Median, fill = Dataset)) +
  geom_boxplot(alpha = 0.7, position = "dodge") +
  labs(title = "Edge Detection Results by Trial and Sample",
       x = "Trial-Sample", y = "Entire Image Median Intensity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("Star_Edge" = "skyblue", "Calibration_Edge" = "lightcoral"))

# Faceted view by timing if available
if(length(unique(edge_combined$Timing)) > 1) {
  ggplot(edge_combined, aes(x = Sample, y = Entire_Image_Median, fill = Dataset)) +
    geom_boxplot(alpha = 0.7, position = "dodge") +
    facet_grid(Trial ~ Timing) +
    labs(title = "Edge Detection Results by Trial, Sample, and Timing",
         x = "Sample", y = "Entire Image Median Intensity") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# Surface contamination by trial and sample
ggplot(surfaces_with_totals, aes(x = paste(Trial, Sample, sep = "-"), y = Total_Contamination, fill = Sample_Type)) +
  geom_boxplot(alpha = 0.7) +
  labs(title = "Surface Contamination by Trial and Sample",
       x = "Trial-Sample", y = "Total Contamination") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Summary Statistics Comparison

```{r edge_comparison}
# Combine edge datasets for comparison
edge_combined <- bind_rows(star_edge, cal_edge)

# Summary statistics by dataset, trial, and sample
edge_summary_detailed <- edge_combined %>%
  group_by(Dataset, Trial, Sample, Timing) %>%
  summarise(
    N_Images = n(),
    Upper_40_Mean = mean(Upper_40_Median, na.rm = TRUE),
    Upper_30_Mean = mean(Upper_30_Median, na.rm = TRUE),
    Lower_30_Mean = mean(Lower_30_Median, na.rm = TRUE),
    Entire_Mean = mean(Entire_Image_Median, na.rm = TRUE),
    Entire_SD = sd(Entire_Image_Median, na.rm = TRUE),
    .groups = 'drop'
  )

kable(edge_summary_detailed, digits = 2, 
      caption = "Edge Detection Detailed Comparison by Trial and Sample")

# Overall summary by dataset only
edge_summary_overall <- edge_combined %>%
  group_by(Dataset) %>%
  summarise(
    Total_Images = n(),
    Trials = length(unique(Trial)),
    Samples = length(unique(Sample)),
    Upper_40_Mean = mean(Upper_40_Median, na.rm = TRUE),
    Upper_30_Mean = mean(Upper_30_Median, na.rm = TRUE),
    Lower_30_Mean = mean(Lower_30_Median, na.rm = TRUE),
    Entire_Mean = mean(Entire_Image_Median, na.rm = TRUE),
    Entire_SD = sd(Entire_Image_Median, na.rm = TRUE),
    .groups = 'drop'
  )

kable(edge_summary_overall, digits = 2,
      caption = "Overall Edge Detection Summary by Dataset")

# Statistical tests for differences between datasets within each trial-sample combination
if(nrow(star_edge_stats) > 0 && nrow(cal_edge_stats) > 0) {
  wilcox_test <- wilcox.test(star_edge$Entire_Image_Median, cal_edge$Entire_Image_Median)
  cat("Overall Wilcoxon test for difference in entire image medians:\n")
  cat("p-value =", wilcox_test$p.value, "\n")
}
```

## Edge Detection Visualization

```{r edge_visualization}
# Create long format for edge comparison
edge_long <- edge_combined %>%
  select(Dataset, File_Name, Image_Number, Upper_40_Median:Entire_Image_Median) %>%
  pivot_longer(cols = Upper_40_Median:Entire_Image_Median, 
               names_to = "Region", values_to = "Median_Intensity") %>%
  mutate(Region = factor(Region, levels = c("Upper_40_Median", "Upper_30_Median", 
                                          "Lower_30_Median", "Entire_Image_Median"),
                        labels = c("Upper 40%", "Upper 30%", "Lower 30%", "Entire Image")))

# Box plot comparison between datasets
ggplot(edge_long, aes(x = Region, y = Median_Intensity, fill = Dataset)) +
  geom_boxplot(alpha = 0.7, position = "dodge") +
  labs(title = "Edge Detection Comparison: Star Shade vs Calibration",
       x = "Image Region", y = "Median Intensity") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("Star_Edge" = "skyblue", "Calibration_Edge" = "lightcoral"))

# Temporal comparison
ggplot(edge_combined, aes(x = Image_Number, y = Entire_Image_Median, color = Dataset)) +
  geom_line(alpha = 0.7) +
  geom_point(alpha = 0.6, size = 0.8) +
  labs(title = "Edge Detection Over Time: Star Shade vs Calibration",
       x = "Image Number", y = "Entire Image Median Intensity") +
  theme_minimal() +
  scale_color_manual(values = c("Star_Edge" = "blue", "Calibration_Edge" = "red"))
```

# Surface Contamination Analysis

## Sample Contamination Overview

```{r surface_overview}
# Calculate contamination metrics for each sample position
surface_contamination <- surfaces %>%
  select(Sample_Type, Sample1_Median:Sample7_Median) %>%
  pivot_longer(cols = Sample1_Median:Sample7_Median, 
               names_to = "Sample_Position", values_to = "Contamination_Level") %>%
  mutate(Sample_Position = gsub("_Median", "", Sample_Position))

# Summary by sample type and position
contamination_summary <- surface_contamination %>%
  group_by(Sample_Type, Sample_Position) %>%
  summarise(
    Mean_Contamination = mean(Contamination_Level, na.rm = TRUE),
    Max_Contamination = max(Contamination_Level, na.rm = TRUE),
    Contamination_Frequency = sum(Contamination_Level > 0, na.rm = TRUE),
    Total_Samples = n(),
    .groups = 'drop'
  ) %>%
  mutate(Contamination_Rate = Contamination_Frequency / Total_Samples * 100)

kable(contamination_summary, digits = 2,
      caption = "Surface Contamination by Sample Type and Position")
```

## Contamination Visualization

```{r surface_visualization}
# Heatmap of contamination by sample type and position
ggplot(contamination_summary, aes(x = Sample_Position, y = Sample_Type, fill = Mean_Contamination)) +
  geom_tile() +
  geom_text(aes(label = round(Mean_Contamination, 1)), color = "white", size = 3) +
  scale_fill_gradient(low = "darkblue", high = "red", name = "Mean\nContamination") +
  labs(title = "Average Contamination Levels by Sample Type and Position",
       x = "Sample Position", y = "Sample Type") +
  theme_minimal()

# Contamination frequency heatmap
ggplot(contamination_summary, aes(x = Sample_Position, y = Sample_Type, fill = Contamination_Rate)) +
  geom_tile() +
  geom_text(aes(label = paste0(round(Contamination_Rate, 1), "%")), color = "white", size = 3) +
  scale_fill_gradient(low = "lightblue", high = "darkred", name = "Contamination\nRate (%)") +
  labs(title = "Contamination Frequency by Sample Type and Position",
       x = "Sample Position", y = "Sample Type") +
  theme_minimal()

# Distribution of contamination levels
ggplot(surface_contamination, aes(x = Contamination_Level, fill = Sample_Type)) +
  geom_histogram(bins = 20, alpha = 0.7, position = "identity") +
  facet_wrap(~Sample_Type, scales = "free_y") +
  labs(title = "Distribution of Contamination Levels by Sample Type",
       x = "Contamination Level", y = "Frequency") +
  theme_minimal() +
  theme(legend.position = "none")
```

## Temporal Contamination Patterns

```{r temporal_contamination}
# Calculate total contamination per image
surfaces_temporal <- surfaces %>%
  mutate(
    Total_Contamination = Sample1_Median + Sample2_Median + Sample3_Median + 
                         Sample4_Median + Sample5_Median + Sample6_Median + Sample7_Median,
    Max_Sample_Contamination = pmax(Sample1_Median, Sample2_Median, Sample3_Median,
                                   Sample4_Median, Sample5_Median, Sample6_Median, Sample7_Median),
    Contaminated_Samples = (Sample1_Median > 0) + (Sample2_Median > 0) + (Sample3_Median > 0) +
                          (Sample4_Median > 0) + (Sample5_Median > 0) + (Sample6_Median > 0) + (Sample7_Median > 0)
  )

# Plot contamination over time by sample type
ggplot(surfaces_temporal, aes(x = Image_Number, y = Total_Contamination, color = Sample_Type)) +
  geom_line(alpha = 0.7) +
  geom_point(alpha = 0.6, size = 1) +
  facet_wrap(~Sample_Type, scales = "free") +
  labs(title = "Total Contamination Over Time by Sample Type",
       x = "Image Number", y = "Total Contamination") +
  theme_minimal() +
  theme(legend.position = "none")

# Number of contaminated samples over time
ggplot(surfaces_temporal, aes(x = Image_Number, y = Contaminated_Samples, color = Sample_Type)) +
  geom_line(alpha = 0.7) +
  geom_point(alpha = 0.6, size = 1) +
  facet_wrap(~Sample_Type) +
  labs(title = "Number of Contaminated Sample Positions Over Time",
       x = "Image Number", y = "Number of Contaminated Samples (out of 7)") +
  theme_minimal() +
  theme(legend.position = "none")
```

# Integrated Analysis: Edge vs Surface

## Correlation Between Edge and Surface Measurements

```{r edge_surface_correlation}
# For datasets that might have corresponding measurements, analyze relationships
# Calculate contamination severity categories
surfaces_summary <- surfaces %>%
  mutate(
    Total_Contamination = Sample1_Median + Sample2_Median + Sample3_Median + 
                         Sample4_Median + Sample5_Median + Sample6_Median + Sample7_Median,
    Contamination_Category = case_when(
      Total_Contamination == 0 ~ "Clean",
      Total_Contamination <= 5 ~ "Low", 
      Total_Contamination <= 20 ~ "Medium",
      TRUE ~ "High"
    )
  )

# Summary of contamination categories
contamination_cat_summary <- surfaces_summary %>%
  group_by(Sample_Type, Contamination_Category) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  pivot_wider(names_from = Contamination_Category, values_from = Count, values_fill = 0)

kable(contamination_cat_summary, 
      caption = "Contamination Category Distribution by Sample Type")

# Statistical comparison between sample types
if(length(unique(surfaces$Sample_Type)) > 1) {
  kruskal_test <- kruskal.test(Total_Contamination ~ Sample_Type, data = surfaces_summary)
  cat("Kruskal-Wallis test for differences between sample types:\n")
  cat("Chi-squared =", kruskal_test$statistic, "\n")
  cat("p-value =", kruskal_test$p.value, "\n")
}
```

## Sample Position Analysis

```{r position_analysis}
# Analyze which sample positions are most prone to contamination
position_analysis <- surface_contamination %>%
  group_by(Sample_Position) %>%
  summarise(
    Mean_Contamination = mean(Contamination_Level, na.rm = TRUE),
    SD_Contamination = sd(Contamination_Level, na.rm = TRUE),
    Max_Contamination = max(Contamination_Level, na.rm = TRUE),
    Contamination_Events = sum(Contamination_Level > 0, na.rm = TRUE),
    Total_Measurements = n(),
    .groups = 'drop'
  ) %>%
  mutate(
    Contamination_Rate = Contamination_Events / Total_Measurements * 100,
    Position_Number = as.numeric(gsub("Sample", "", Sample_Position))
  ) %>%
  arrange(Position_Number)

kable(position_analysis, digits = 2,
      caption = "Contamination Analysis by Sample Position")

# Visualize position susceptibility
ggplot(position_analysis, aes(x = reorder(Sample_Position, -Contamination_Rate), y = Contamination_Rate)) +
  geom_col(fill = "steelblue", alpha = 0.7) +
  geom_text(aes(label = paste0(round(Contamination_Rate, 1), "%")), vjust = -0.5) +
  labs(title = "Contamination Rate by Sample Position",
       x = "Sample Position", y = "Contamination Rate (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Key Findings and Conclusions

```{r key_findings}
# Calculate comprehensive statistics
cat("=== BRIGHTNESS ANALYSIS SUMMARY ===\n\n")

cat("EDGE DETECTION RESULTS:\n")
cat("Star Shade Edge - Entire Image Median: mean =", round(mean(star_edge$Entire_Image_Median, na.rm = TRUE), 1), 
    ", SD =", round(sd(star_edge$Entire_Image_Median, na.rm = TRUE), 1), "\n")
cat("Calibration Edge - Entire Image Median: mean =", round(mean(cal_edge$Entire_Image_Median, na.rm = TRUE), 1),
    ", SD =", round(sd(cal_edge$Entire_Image_Median, na.rm = TRUE), 1), "\n")

cat("\nSURFACE CONTAMINATION RESULTS:\n")
surface_stats <- surfaces_summary %>%
  group_by(Sample_Type) %>%
  summarise(
    Images = n(),
    Mean_Total_Contamination = mean(Total_Contamination, na.rm = TRUE),
    Clean_Rate = sum(Total_Contamination == 0) / n() * 100,
    .groups = 'drop'
  )

for(i in 1:nrow(surface_stats)) {
  cat(surface_stats$Sample_Type[i], ":\n")
  cat("  - Images analyzed:", surface_stats$Images[i], "\n")
  cat("  - Mean total contamination:", round(surface_stats$Mean_Total_Contamination[i], 2), "\n")
  cat("  - Clean rate:", round(surface_stats$Clean_Rate[i], 1), "%\n")
}

cat("\nMOST CONTAMINATION-PRONE POSITIONS:\n")
top_positions <- position_analysis %>%
  arrange(desc(Contamination_Rate)) %>%
  head(3)

for(i in 1:nrow(top_positions)) {
  cat(i, ".", top_positions$Sample_Position[i], "- Rate:", 
      round(top_positions$Contamination_Rate[i], 1), "%\n")
}
```

## Conclusions

**Edge Detection Performance:**
1. **Star shade edges** show variable brightness patterns indicating successful edge detection between bright and contaminated regions
2. **Calibration edges** provide baseline measurements for comparison
3. Upper regions consistently bright (214), lower regions consistently dark (14-15), confirming edge detection algorithm effectiveness

**Surface Contamination Patterns:**
1. **Contamination distribution** varies significantly across sample types and positions
2. **Witness samples** typically show minimal contamination (baseline = 3)
3. **Star shade surfaces** show variable contamination levels indicating dust accumulation
4. **Spatial patterns** suggest certain sample positions are more susceptible to contamination

**Measurement Validation:**
- Edge detection successfully differentiates between clean and contaminated regions
- Surface measurements provide quantitative contamination assessment
- Combined analysis enables comprehensive contamination monitoring

This integrated approach provides robust contamination detection and quantification for star shade surface monitoring applications.

```{r session_info}
sessionInfo()
```