---
title: "Updated Comprehensive Surface Comparison Analysis: After vs Before Next Trial"
author: "Data Analyst"
date: "`r Sys.Date()`"
output: 
  pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(knitr)
library(rstatix)
library(ggpubr)
library(gridExtra)
library(data.table)
library(cowplot)
library(ggplot2)
library(dplyr)
library(grid)
```

## 1. Data Loading and Preprocessing

We'll modify our data loading function to include trials 1-8, then create comparison pairs.

```{r data_loading, echo = FALSE, warning = FALSE, message = FALSE}
# Function to load and process data (either before or after)
load_data <- function(condition) {
  # Function to load and process particle data for a single sample
  load_particle_data <- function(trial_number, sample_number, cond) {
    tryCatch({
      particle_path <- sprintf("~/Desktop/College/Research/Dust_Contamination/Trials/Data/Surfaces/%sTr%dSa%dSurf/Particles_%sTr%dSa%dSurf.csv", 
                               cond, trial_number, sample_number, cond, trial_number, sample_number)
      
      particle_data <- read_csv(particle_path)
      particle_data <- particle_data[, 2:(ncol(particle_data)-2)]
      particle_data$Diameter <- sqrt(particle_data$Area / pi) * 2
      particle_data$Sample <- sample_number
      particle_data$Trial <- trial_number
      
      return(particle_data)
    }, error = function(e) {
      message(sprintf("Error loading particle data for %s Trial %d, Sample %d: %s", cond, trial_number, sample_number, e$message))
      return(NULL)
    })
  }

  # Function to load and process summary data for a single sample
  load_summary_data <- function(trial_number, sample_number, cond) {
    tryCatch({
      summary_path <- sprintf("~/Desktop/College/Research/Dust_Contamination/Trials/Data/Surfaces/%sTr%dSa%dSurf/Summary_%sTr%dSa%dSurf.csv", 
                              cond, trial_number, sample_number, cond, trial_number, sample_number)
      
      summary_data <- read_csv(summary_path, col_select = c(1, 2, 3, 5))
      names(summary_data) <- make.names(names(summary_data))
      summary_data <- summary_data %>% 
        drop_na() %>%
        mutate(
          Sample = sample_number,
          Trial = trial_number,
          Slice_Number = as.integer(substr(Slice, nchar(Slice)-6, nchar(Slice)-4))
        )
      
      return(summary_data)
    }, error = function(e) {
      message(sprintf("Error loading summary data for %s Trial %d, Sample %d: %s", cond, trial_number, sample_number, e$message))
      return(NULL)
    })
  }

  # Load data for trials 1-8 and 5 samples in each trial
  all_particle_data <- map2(rep(1:8, each = 5), rep(1:5, times = 8), 
                            ~load_particle_data(.x, .y, condition))
  all_summary_data <- map2(rep(1:8, each = 5), rep(1:5, times = 8), 
                           ~load_summary_data(.x, .y, condition))

  # Combine all data into single data frames
  combined_particle_data <- bind_rows(all_particle_data)
  combined_summary_data <- bind_rows(all_summary_data)

  # Add cleaning method information
  add_cleaning_method <- function(data) {
    data %>%
      mutate(Cleaning_Method = case_when(
        Trial %in% c(1, 2, 3) ~ "IPA rinse",
        Trial %in% c(4, 5) ~ "Drag and wipe",
        Trial %in% c(6, 7) ~ "First contact",
        Trial == 8 ~ "First contact & Drag and wipe",
        TRUE ~ NA_character_
      ))
  }

  combined_particle_data <- add_cleaning_method(combined_particle_data)
  combined_summary_data <- add_cleaning_method(combined_summary_data)

  # Calculate total imaged area for each trial
  image_size <- 600 * 450 # microns^2
  trial_areas <- combined_summary_data %>%
    group_by(Sample, Trial, Cleaning_Method) %>%
    summarise(
      Total_Images = n(),
      Total_Area = Total_Images * image_size * 1e-12, # Convert to m^2
      .groups = "drop"
    )

  # Normalize particle counts to represent 0.1 m^2
  normalization_factor <- 0.1 / trial_areas$Total_Area
  combined_particle_data <- combined_particle_data %>%
    left_join(trial_areas, by = c("Trial", "Cleaning_Method")) %>%
    mutate(Normalized_Count = normalization_factor[Trial])

  list(particle_data = combined_particle_data, 
       summary_data = combined_summary_data, 
       trial_areas = trial_areas)
}

# Load before and after data
before_data <- load_data("Bef")
after_data <- load_data("Aft")

# Create comparison pairs
create_comparison_pairs <- function(after_data, before_data) {
  comparison_pairs <- list()
  for (i in 1:7) {
    comparison_pairs[[i]] <- list(
      after = after_data$particle_data %>% filter(Trial == i),
      before = before_data$particle_data %>% filter(Trial == i + 1),
      after_summary = after_data$summary_data %>% filter(Trial == i),
      before_summary = before_data$summary_data %>% filter(Trial == i + 1)
    )
  }
  return(comparison_pairs)
}

comparison_pairs <- create_comparison_pairs(after_data, before_data)
```

## 3. Cleaning Method Effectiveness

Let's examine how the effectiveness of each cleaning method changed between trials.

```{r cleaning_method_effectiveness}
# Function to calculate particle statistics
get_particle_stats <- function(data) {
  data %>%
    group_by(Cleaning_Method) %>%
    summarise(
      Total_Particles = n() * mean(Normalized_Count),
      Mean_Diameter = mean(Diameter),
      Median_Diameter = median(Diameter),
      SD_Diameter = sd(Diameter)
    )
}

# Calculate stats for each comparison pair
stats_list <- lapply(1:7, function(i) {
  after_stats <- get_particle_stats(comparison_pairs[[i]]$after)
  before_stats <- get_particle_stats(comparison_pairs[[i]]$before)
  
  bind_rows(
    mutate(after_stats, Condition = paste("After Trial", i)),
    mutate(before_stats, Condition = paste("Before Trial", i+1))
  )
})

# Display results
for (i in 1:7) {
  print(kable(stats_list[[i]], caption = paste("Particle Statistics: After Trial", i, "vs Before Trial", i+1)))
  
  # Visualize Total Particles
  print(ggplot(stats_list[[i]], aes(x = Cleaning_Method, y = Total_Particles, fill = Condition)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = paste("Total Particles: After Trial", i, "vs Before Trial", i+1),
         y = "Total Particles (Normalized to 0.1 m^2)") +
    theme_minimal())
  
  # Visualize Mean Diameter
  print(ggplot(stats_list[[i]], aes(x = Cleaning_Method, y = Mean_Diameter, fill = Condition)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = paste("Mean Particle Diameter: After Trial", i, "vs Before Trial", i+1),
         y = "Mean Diameter (microns)") +
    theme_minimal())
}
```

## 4. Slice-by-Slice Analysis

Now, let's perform a slice-by-slice analysis to compare the particle distribution across different slices for each pair of trials.

```{r slice_analysis}
# Function to prepare slice data
prepare_slice_data <- function(data, condition) {
  data %>%
    group_by(Trial, Sample, Slice_Number, Cleaning_Method) %>%
    summarise(Total_Area = sum(Total.Area), .groups = "drop") %>%
    mutate(Condition = condition)
}

# Create and display heatmaps for each comparison pair
for (i in 1:7) {
  after_slice_data <- prepare_slice_data(comparison_pairs[[i]]$after_summary, paste("After Trial", i))
  before_slice_data <- prepare_slice_data(comparison_pairs[[i]]$before_summary, paste("Before Trial", i+1))
  
  combined_slice_data <- bind_rows(after_slice_data, before_slice_data)
  
  # Create heatmaps for each sample
  sample_heatmaps <- lapply(1:5, function(sample) {
    heatmap_data <- combined_slice_data %>%
      filter(Sample == sample)
    
    ggplot(heatmap_data, aes(x = Slice_Number, y = Condition, fill = Total_Area)) +
      geom_tile() +
      scale_fill_gradient(low = "white", high = "red") +
      scale_x_continuous(breaks = seq(0, 500, by = 100)) +
      labs(title = paste("Sample", sample, ": After Trial", i, "vs Before Trial", i+1),
           x = "Slice Number", y = "Condition", fill = "Total Area") +
      theme_minimal()
  })
  
  # Display heatmaps
  do.call(grid.arrange, c(sample_heatmaps, ncol = 2))
}
```


## 6. Comparison to IEST Standards

Finally, let's compare how the adherence to IEST standards changed between each pair of trials.
```{r iest_comparison}
# Define IEST standard parameters
slope <- -0.926

# Function to calculate normalized counts
get_normalized_counts <- function(data, trial_area) {
  sapply(diameter_thresholds, function(x) {
    sum(data$Diameter > x) * (0.1 / trial_area)
  })
}

# Calculate and plot IEST comparisons for each pair
for (i in 1:7) {
  after_trial_area <- after_data$trial_areas$Total_Area[after_data$trial_areas$Trial == i]
  before_trial_area <- before_data$trial_areas$Total_Area[before_data$trial_areas$Trial == i+1]
  
  after_counts <- get_normalized_counts(comparison_pairs[[i]]$after, after_trial_area)
  before_counts <- get_normalized_counts(comparison_pairs[[i]]$before, before_trial_area)
  
  counts_data <- data.frame(
    Diameter = log10(diameter_thresholds)^2,
    After_Count = after_counts,
    Before_Count = before_counts
  )
  
  counts_data <- counts_data %>% 
      filter(After_Count > 0) %>% 
    filter(Before_Count > 0)
  
  # Calculate best fit lines
  after_fit <- lm(log10(After_Count) ~ Diameter, data = counts_data)
  before_fit <- lm(log10(Before_Count) ~ Diameter, data = counts_data)
  
  best_fit_lines <- data.frame(
    Condition = c(paste("After Trial", i), paste("Before Trial", i+1)),
    Slope = c(coef(after_fit)[2], coef(before_fit)[2]),
    Intercept = c(coef(after_fit)[1], coef(before_fit)[1])
  ) %>%
    mutate(PCL = 10^(sqrt(abs(Intercept / Slope))))
  
  counts_data <- data.frame(
    Diameter = log10(diameter_thresholds)^2,
    After_Count = after_counts,
    Before_Count = before_counts
  ) %>%
    pivot_longer(cols = c(After_Count, Before_Count), 
                 names_to = "Condition", 
                 values_to = "Count")
  
  # Plot comparison to IEST standards
  iest_plot <- ggplot(counts_data, aes(x = Diameter, y = Count, color = Condition)) +
    geom_line() +
    geom_abline(data = best_fit_lines, 
                aes(slope = Slope, intercept = Intercept, color = Condition),
                linetype = "dashed") +
    scale_y_log10() +
    labs(x = "Diameter (log(microns)^2)", 
         y = "Count of Particles (log scale)",
         title = paste("IEST Comparison: After Trial", i, "vs Before Trial", i+1),
         subtitle = "Solid lines represent observed data, dashed lines represent fitted IEST standards") +
    theme_minimal()
  
  print(iest_plot)
  
  # Display PCL and Slope statistics
  print(kable(best_fit_lines, caption = paste("PCL and Slope Statistics: After Trial", i, "vs Before Trial", i+1)))
}
```

## Comprehensive Sample Comparison

To provide a holistic view of how particle distributions change across all samples and trials, we've created a comprehensive comparison graph. This graph displays the normalized particle counts for each sample in both the "after" and "before" conditions for all trials.

```{r comprehensive_comparison, fig.width=12, fig.height=8}

# Function to prepare data for a single trial pair
prepare_trial_pair_data <- function(pair, trial_num) {
  after_data <- pair$after %>%
    group_by(Diameter) %>%
    summarise(
      Counts = list(sapply(diameter_thresholds, function(x) sum(Diameter > x) * unique(Normalized_Count))),
      .groups = "drop"
    )
  
  before_data <- pair$before %>%
    group_by(Diameter) %>%
    summarise(
      Counts = list(sapply(diameter_thresholds, function(x) sum(Diameter > x) * unique(Normalized_Count))),
      .groups = "drop"
    )
  
  after_data <- after_data %>%
    unnest(Counts) %>%
    mutate(
      Diameter = rep(log10(diameter_thresholds)^2, n_distinct(Sample)),
      Condition = paste0("After Trial ", trial_num),
      Sample = paste0("Sample ", Sample)
    )
  
  before_data <- before_data %>%
    unnest(Counts) %>%
    mutate(
      Diameter = rep(log10(diameter_thresholds)^2, n_distinct(Sample)),
      Condition = paste0("Before Trial ", trial_num + 1),
      Sample = paste0("Sample ", Sample)
    )
  
  bind_rows(after_data, before_data)
}

# Prepare data for all trial pairs
all_pairs_data <- map2_dfr(comparison_pairs, 1:7, prepare_trial_pair_data)

# Create the comprehensive comparison plot
ggplot(all_pairs_data, aes(x = Diameter, y = Counts, color = interaction(Condition, Sample))) +
  geom_line() +
  scale_y_log10() +
  labs(x = "Diameter (log(microns)^2)", 
       y = "Normalized Count of Particles (log scale)",
       title = "Comprehensive Comparison of Particle Distributions Across All Samples and Trials",
       color = "Condition and Sample") +
  theme_minimal() +
  theme(legend.position = "right",
        legend.text = element_text(size = 6),
        legend.title = element_text(size = 8)) +
  guides(color = guide_legend(ncol = 2))
```

```{r}
# Function to prepare averaged data for a single trial pair
prepare_averaged_trial_pair_data <- function(pair, trial_num) {
  after_data <- pair$after %>%
    group_by(Diameter, Sample) %>%
    summarise(
      Avg_Count = mean(Normalized_Count * (Diameter > Diameter)),
      .groups = "drop"
    ) %>%
    mutate(
      Condition = paste0("After Trial ", trial_num),
      Diameter = log10(Diameter)^2
    )
  
  before_data <- pair$before %>%
    group_by(Diameter, Sample) %>%
    summarise(
      Avg_Count = mean(Normalized_Count * (Diameter > Diameter)),
      .groups = "drop"
    ) %>%
    mutate(
      Condition = paste0("Before Trial ", trial_num + 1),
      Diameter = log10(Diameter)^2
    )
  
  bind_rows(after_data, before_data)
}

# Prepare averaged data for all trial pairs
all_averaged_data <- map2_dfr(comparison_pairs, 1:7, prepare_averaged_trial_pair_data)

# Create the average trial comparison plot
ggplot(all_averaged_data, aes(x = Diameter, y = Avg_Count, color = Condition)) +
  geom_line() +
  scale_y_log10() +
  labs(x = "Diameter (log(microns)^2)", 
       y = "Average Normalized Count of Particles (log scale)",
       title = "Average Particle Distributions Across Trials",
       color = "Condition") +
  theme_minimal() +
  theme(legend.position = "right",
        legend.text = element_text(size = 8),
        legend.title = element_text(size = 10)) +
  scale_color_manual(values = c(
    "After Trial 1" = "#FF0000", "Before Trial 2" = "#FF9999",
    "After Trial 2" = "#00FF00", "Before Trial 3" = "#99FF99",
    "After Trial 3" = "#0000FF", "Before Trial 4" = "#9999FF",
    "After Trial 4" = "#FFFF00", "Before Trial 5" = "#FFFF99",
    "After Trial 5" = "#FF00FF", "Before Trial 6" = "#FF99FF",
    "After Trial 6" = "#00FFFF", "Before Trial 7" = "#99FFFF",
    "After Trial 7" = "#800000", "Before Trial 8" = "#FF8080"
  ))
```



```{r}
# Define IEST standard parameters
slope <- -0.926

# Function to calculate normalized counts for each sample
get_normalized_counts <- function(data, trial_sample_areas) {
  if (is.null(data) || nrow(data) == 0) {
    warning("Input data is NULL or empty")
    return(NULL)
  }
  if (is.null(trial_sample_areas) || nrow(trial_sample_areas) == 0) {
    warning("Trial sample areas data is NULL or empty")
    return(NULL)
  }
  
  tryCatch({
    result <- data %>%
      left_join(trial_sample_areas, by = c("Trial", "Sample", "Cleaning_Method")) %>%
      group_by(Trial, Sample) %>%
      summarise(
        Counts = list(sapply(diameter_thresholds, function(x) sum(Diameter > x) * (0.1 / first(Total_Area)))),
        .groups = "drop"
      ) %>%
      unnest(Counts) %>%
      group_by(Trial) %>%
      summarise(
        Avg_Count = mean(Counts),
        .groups = "drop"
      )
    return(result)
  }, error = function(e) {
    warning(paste("Error in get_normalized_counts:", e$message))
    return(NULL)
  })
}

# Calculate and plot IEST comparisons for each pair
for (i in 1:7) {
  cat(paste("\nProcessing comparison pair:", i, "\n"))
  
  # Print diagnostic information
  cat("After data summary:\n")
  print(summary(comparison_pairs[[i]]$after))
  cat("\nBefore data summary:\n")
  print(summary(comparison_pairs[[i]]$before))
  cat("\nAfter areas summary:\n")
  print(summary(comparison_pairs[[i]]$after_areas))
  cat("\nBefore areas summary:\n")
  print(summary(comparison_pairs[[i]]$before_areas))
  
  after_counts <- get_normalized_counts(comparison_pairs[[i]]$after, comparison_pairs[[i]]$after_areas)
  before_counts <- get_normalized_counts(comparison_pairs[[i]]$before, comparison_pairs[[i]]$before_areas)
  
  if (is.null(after_counts) || is.null(before_counts)) {
    warning(paste("Skipping comparison pair", i, "due to NULL counts"))
    next
  }
  
  counts_data <- data.frame(
    Diameter = log10(diameter_thresholds)^2,
    After_Count = after_counts$Avg_Count,
    Before_Count = before_counts$Avg_Count
  )
  
  counts_data <- counts_data %>% 
    filter(After_Count > 0, Before_Count > 0)
  
  if (nrow(counts_data) == 0) {
    warning(paste("No valid data for comparison pair", i))
    next
  }
  
  # Calculate best fit lines
  after_fit <- lm(log10(After_Count) ~ Diameter, data = counts_data)
  before_fit <- lm(log10(Before_Count) ~ Diameter, data = counts_data)
  
  best_fit_lines <- data.frame(
    Condition = c(paste("After Trial", i), paste("Before Trial", i+1)),
    Slope = c(coef(after_fit)[2], coef(before_fit)[2]),
    Intercept = c(coef(after_fit)[1], coef(before_fit)[1])
  ) %>%
    mutate(PCL = 10^(sqrt(abs(Intercept / Slope))))
  
  # Prepare data for plotting
  plot_data <- counts_data %>%
    pivot_longer(cols = c(After_Count, Before_Count), 
                 names_to = "Condition", 
                 values_to = "Count") %>%
    mutate(Condition = ifelse(Condition == "After_Count", 
                              paste("After Trial", i), 
                              paste("Before Trial", i+1)))
  
  # Plot comparison to IEST standards
  iest_plot <- ggplot(plot_data, aes(x = Diameter, y = Count, color = Condition)) +
    geom_point() +
    geom_line() +
    geom_abline(data = best_fit_lines, 
                aes(slope = Slope, intercept = Intercept, color = Condition),
                linetype = "dashed") +
    scale_y_log10() +
    labs(x = "Diameter (log(microns)^2)", 
         y = "Average Normalized Count of Particles (log scale)",
         title = paste("IEST Comparison: After Trial", i, "vs Before Trial", i+1),
         subtitle = "Points and solid lines represent observed data, dashed lines represent fitted IEST standards") +
    theme_minimal()
  
  print(iest_plot)
  
  # Display PCL and Slope statistics
  print(kable(best_fit_lines, caption = paste("PCL and Slope Statistics: After Trial", i, "vs Before Trial", i+1)))
}
```
```
```

